Bare Demo of IEEEtran.cls\\ for IEEE Conferences

### 2016-12-01 paper.tex
Better description of parameters used.

Learning to program is often regarded as a difficult task, courses are 
problem-solving intensive and often students feel frustrated when doing programming
exercises. Instructors aim to design learning activities that result
challenging to students, but not to the extreme as to frustrate them. When
selecting an appropiate programing excercise, experienced instructors gauge a
student’s affective state and skills to then assign an activity with the  appropriate level
of difficulty. This work is focused on the prediction of the affective states
of programmers with different levels of expertise when learning a new
programming language. For this, an interactive web-based programming platform
is proposed. The platform is designed to collect data from the students'
interaction for data analysis. Current work is focused on the prediction of
affective states using non-obtrusive sensors. Specifically, the proposed method uses data
gathered from both keyboard and mouse dynamics. The data set was generated
as students learned the basics of Python programming. Feature vectors 
were generated with 39 features representing the dynamics of the interaction. 
In total 149 feature vectors were generated and used as input of four different
classification algorithms: Na\"ive Bayes, J-48, k-NN, and Artificial Neural
Networks. These classification models are then compared by their performances
when classifying learners into five different affective states: boredom,
frustration, distraction, relaxation and engagement. Each of the classification
models achieved an accuracy of around 75\% with a kappa of around 0.45, and, in
average, J-48 was the better algorithm with an accuracy of 75.97\%. Results show
that the data gathered from the non-obtrusive sensors can successfully be used
as another input to classification models in order to predict an individual's
affective states.

@@

The structure of this work is organized as follows: next Section 
presents a series of works related to the proposed method in this paper; Section
\ref{sec:method} describes the proposed method for the recognition of affective
states in a web based learning environment for the teaching of programming
languages; next in Section \ref{sec:exp} we explain the experimental evaluation of the
proposed method. % There is no results section so far
Finally, we draw some conclusions about our methodology and
experimients.

@@

In order to determine what affective states a student was experiencing, the
Experience Sampling Method (ESM) \cite{kubey1996experience} was used .
After the students successfully solve a programming exercise, they are presented
with an ESM survey that asks what they were feeling during their solving of the
exercise. A very brief description is given about what to do in this survey,
followed by statements the students need to answer according to how they were
feeling. As an example, the statement “I was feeling frustrated” is presented,
and a student needs to answer either “Strongly agree,” “Agree,” “Neutral,”
“Disagree,” and “Strongly Disagree.”

@@

Classifiers where implemented using the software Rapid Miner using a feature
selection module % which one? - JJ
as part of the process. A genetic algorithm was used in the
user selection module. % Which one? You really have to be _way_ more
                       % specific in this conference - JJ
In the end, the subset for the frustration classifier
consists of 11 features, and the subset for the boredom classifier consists of
13 features. The performance of the classifiers is shown in Table \ref{tab_performance}.

%Start with a reference to the Table (and check numbering) - JJ
The results obtained with this method are satisfactory, considering a free-text
was used. % don't get this. What free text? A program? - JJ
The accuracies and kappa coefficients obtained are close to what is
usually obtained in fixed-text methods, for example, the results presented in
Epp, Lippold y Mandryk \cite{epp2011identifying}. %A reference!!! - JJ  
												 %Added	
 In the case of the kappa coefficients, it is
usual to see values below 0.2 in methods involving free-text. In this case, some
values of kappa were close to 0.5, is important to consider the values of kappa
in these results because the class distribution is not uniformly distributed. As
it is observed in many works in affective recognition, decision trees normally
produce competitive accuracy. In this case the J-48 classifier obtained an
accuracy of 80.48\%, which is marginally better than the others, and an accepted kappa
statistic for the kind of problem. The artificial neural network % Which one? - JJ
give the highest
accuracy in the classification of frustration.

@@

Future work will focus on extracting other features from the learners’ interaction
like behavioral data, as proposed by Blikstein (2011); the data includes the size
of the program and changes in the code. Other authors also propose treating the use
of certain keys as features e.g., the delete key because is manly used for correcting
sentences. Also a multimodal approach is needed.

>>>>

Learning to program is often regarded as a difficult task, courses are 
problem-solving intensive and often students feel frustrated when doing programming
exercises. Instructors aim to design learning activities that result
challenging to students, but not to the extreme as to frustrate them. When
selecting an appropiate programing excercise, experienced instructors gauge a
student’s affective state and skills to then assign an activity with the  appropriate level
of difficulty. This work is focused on the prediction of the affective states
of programmers with different levels of expertise when learning a new
programming language. For this, an interactive web-based programming platform
is proposed. The platform is designed to collect data from the students'
interaction for data analysis. Current work is focused on the prediction of
affective states using non-obtrusive sensors. Specifically,  the aim of this research 
is to evaluate the use of keyboard and mouse dynamics as an appropriate 
sensory input for an affective recognition
system. The proposed method uses a data mining approach in data 
genearated from both keyboard and mouse dynamics of students as they 
work in programming assignments. The data set was generated
as students learned the basics of Python programming. Feature vectors 
were generated with 39 features representing the dynamics of the interaction. 
In total 149 feature vectors were generated and used as input of four different
classification algorithms: Na\"ive Bayes, J-48, k-NN, and Artificial Neural
Networks. These classification models are then compared by their performances
when classifying learners into five different affective states: boredom,
frustration, distraction, relaxation and engagement. Each of the classification
models achieved an accuracy of around 75\% with a kappa of around 0.45, and, in
average, J-48 was the better algorithm with an accuracy of 75.97\%. Results show
that the data gathered from the non-obtrusive sensors can successfully be used
as another input to classification models in order to predict an individual's
affective states.

@@

The structure of this work is organized as follows: next Section 
presents a series of works related to the proposed method in this paper; Section
\ref{sec:method} describes the proposed method for the recognition of affective
states in a web based learning environment for the teaching of programming
languages; next in Section \ref{sec:exp} we explain the experimental evaluation of the
proposed method, and results are discussed in Section \ref{sec:res}  % There is no results section so far
Finally, we draw some conclusions about our methodology and
experimients.

@@

In order to determine what affective states a student was experiencing, the
Experience Sampling Method (ESM) \cite{kubey1996experience} was used.
After the students successfully solve a programming exercise, they were presented
with an ESM survey that asks what they were feeling during their solving of the
exercise. A very brief description is given about what to do in this survey,
followed by statements the students need to answer according to how they were
feeling. As an example, the statement “I was feeling frustrated” is presented,
and a student needs to answer either “Strongly agree,” “Agree,” “Neutral,”
“Disagree,” and “Strongly Disagree.”

@@

As part of the data mining process a genetic algorithm was used for feature
selection with the following parameters: population size 700, 30 generations with
tournament selection, tournament size was 25\% of population size. A uniform crossover 
with 0.5 probability and mutation of 1/number of features, and a minimum feature size of 5.
  % Which one? You really have to be _way_ more
                       % specific in this conference - JJ
In the end, the subset for the frustration classifier
consists of 11 features, and the subset for the boredom classifier consists of
13 features.

The selected classifiers where implemented using Rapid Miner, with the following patameters:

\section{Results}
\label{sec:res} 
In Table \ref{tab_performance} the performance of each of the four classifiers is together 
with the kappa coefficients. A 10-Fold cross validation was used. 
The accuracies and kappa coefficients obtained are close to what is
usually obtained in fixed-text methods, for example, the results presented in
Epp, Lippold y Mandryk \cite{epp2011identifying}. %A reference!!! - JJ  
												 %Added	
In a fixed-text method, subjects are asked to write specific texts.											 
The results obtained with this method are satisfactory, considering that learners
were writing a program, a task comparable with free-text writing, where there
is no restriction on what has to be writen. % don't get this. What free text? A program? - JJ
In the case of the kappa coefficients, it is
usual to see values below 0.2 in methods involving free-text. In this case, some
values of kappa were close to 0.5, is important to consider the values of kappa
in these results because the class distribution is not uniformly distributed. As
it is observed in many works in affective recognition, decision trees normally
produce competitive accuracy. In this case the J-48 classifier obtained an
accuracy of 80.48\%, which is marginally better than the others, and an accepted kappa
statistic for the kind of problem. The artificial neural network % Which one? - JJ
give the highest
accuracy in the classification of frustration.

@@

Future work will focus on extracting other features from the learners’ interaction
like behavioral data, as proposed by Blikstein \cite{blikstein2011using} ; the data includes the size
of the program and changes in the code. Other authors also propose treating the use
of certain keys as features e.g., the delete key because is manly used for correcting
sentences. Also a multimodal approach is needed.
